Algebraic Groups\vspace{-1ex}

### 2019-10-04 Algebraic Geometry/Algebraic-Geometry.tex
Update

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Spec}{Spec}
\let\O\relax
\newcommand*{\O}{\calO}
\newcommand{\A}[1]{\bbA^{#1}}
\let\P\relax
\newcommand{\P}[1]{\bbP^{#1}}

@@

This certainly seems like it should be a regular function (e.g. it is rational and defined everywhere on $\bbR$) but this 
conflicts with the idea that we want to identify $\O_{\A1}(\A1)=k[t]$, but that is clearly not the case here. Think about this.
\subsubsection{Yet another}
Consider the set $R$ of all continuous maps $k\to k$ under the cofinite topology. Someone asked if $R$ is a $k$-algebra. The answer is 
a bit convoluted, but the short answer is no. Specifically if we are using the product topology on $k\times k$, the addition map isn't continuous! 
This also points to the question of what topology is the correct one to use on these things.

@@

The axioms of a space with functions tells us that the \textbf{vanishing locus}
\[Z(f)=\{a|f(a)=0\}\subseteq k^n\]
and furthermore $Z(S)$ should be closed for all $S\subseteq k[x_1,\dots,x_n]$. This leads us to a definition:
\begin{defn}
	A subset $Z\subseteq k^n$ is \textbf{Zariski-closed} if there exists an $S\subseteq k[x_1,\dots,x_n]$
	such that $Z=Z(S)$.
\end{defn}
\begin{lem}
	The Zariski closed sets are the closed sets of a topology (called the \textbf{Zariski Topology}).
\end{lem}
\begin{prf}
	Just do it. Nike. \checked
\end{prf}
\begin{rmk}
	Notice that here the set $\{(a,-a)\}\subseteq k^2$ (the pullback of zero under the addition map) is Zariski closed!
	This fixes the problem we were running into in the third question above.
\end{rmk}
Now since $Z(S)=Z(I_S)$ where $I_S$ is the ideal generated by $S$, it is enough to consider vanishing loci of ideals. Furthermore we have the 
map that extracts the ideal of functions that vanish on a set $Z\subseteq k^n$. There are a ton of great identiies you can prove here. Go to your 
favorite algebra book (e.g. Dummit \& Foote) to see them.

\subsubsection{Functions}
What about functions on these spaces? If we take $f\in k[x_1,\dots,x_n]$ these seem like they should be regular functions $k^n\to k$.
\begin{thm}[(Weak) Nullstellensatz]
	Say $k=\bar k$. Then every maximal ideal $\frakm\lhd k[x_1,\dots,x_n]$ has the form $(x_1-a_1,\dots,x_n-a_n)$.
\end{thm}
\begin{rmk}
	Equivalently, it is the kernel of a $k$-algebra morphis $k[x_1,\dots,x_n]\to k$.
\end{rmk}
\begin{cor}[Nullstellensatz]
	Let $J$ be an ideal of $k[x_1,\dots,x_n]$. Then $I(Z(J))=\sqrt{J}$.
\end{cor}
\begin{prf}
	\textit{Notice this only works when $k$ is uncountable!} Suppose that $\frakm$ is a maximal ideal with residue field 
	$L=k[x_i]/\frakm$. This gives us a surjection of $k[x_1,\dots,x_n]\to L$. Thus $\dim_k L$ is countable!

@@

>>>>

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Spec}{Spec}
\let\O\relax
\newcommand*{\O}{\calO}
\newcommand{\A}[1]{\bbA^{#1}}
\let\P\relax
\newcommand{\P}[1]{\bbP^{#1}}
\newcommand{\eqdef}{\stackrel{\text{def}}{=}}

@@

This certainly seems like it should be a regular function (e.g. it is rational and defined everywhere on $\bbR$) but this 
conflicts with the idea that we want to identify $\O_{\A1}(\A1)=k[t]$, but that is clearly not the case here. Think about this.
\subsubsection{Yet another}\label{subsubsec:third}
Consider the set $R$ of all continuous maps $k\to k$ under the cofinite topology. Someone asked if $R$ is a $k$-algebra. The answer is 
a bit convoluted, but the short answer is no. Specifically if we are using the product topology on $k\times k$, the addition map isn't continuous! 
This also points to the question of what topology is the correct one to use on these things.

@@

The axioms of a space with functions tells us that the \textbf{vanishing locus}
\[Z(f)=\{a|f(a)=0\}\subseteq k^n\]
and furthermore $Z(S)$ should be closed for all $S\subseteq k[x_1,\dots,x_n]$. This leads us to a definition:
\begin{defn}
	A subset $Z\subseteq k^n$ is \textbf{Zariski-closed} if there exists an $S\subseteq k[x_1,\dots,x_n]$
	such that $Z=Z(S)$.
\end{defn}
\begin{lem}
	The Zariski closed sets are the closed sets of a topology (called the \textbf{Zariski Topology}).
\end{lem}
\begin{prf}
	Just do it. Nike. \checked
\end{prf}
\begin{rmk}
	Notice that here the set $\{(a,-a)\}\subseteq k^2$ (the pullback of zero under the addition map) is Zariski closed!
	This fixes the problem we were running into in the third question (sec.~\ref{subsubsec:third}) above.
\end{rmk}
Now since $Z(S)=Z(I_S)$ where $I_S$ is the ideal generated by $S$, it is enough to consider vanishing loci of ideals. Furthermore we have the 
map that extracts the ideal of functions that vanish on a set $Z\subseteq k^n$. There are a ton of great identiies you can prove here. Go to your 
favorite algebra book (e.g. Dummit \& Foote) to see them.

\subsubsection{Functions}
What about functions on these spaces? If we take $f\in k[x_1,\dots,x_n]$ these seem like they should be regular functions $k^n\to k$.
\begin{thm}[(Weak) Nullstellensatz]
	Say $k=\bar k$. Then every maximal ideal $\frakm\lhd k[x_1,\dots,x_n]$ has the form $(x_1-a_1,\dots,x_n-a_n)$.
\end{thm}
\begin{rmk}
	Equivalently, it is the kernel of a $k$-algebra morphis $k[x_1,\dots,x_n]\to k$.
\end{rmk}
\begin{cor}[Nullstellensatz]\label{nullstellensatz}
	Let $J$ be an ideal of $k[x_1,\dots,x_n]$. Then $I(Z(J))=\sqrt{J}$.
\end{cor}
\begin{prf}
	\textit{Notice this only works when $k$ is uncountable!} Suppose that $\frakm$ is a maximal ideal with residue field 
	$L=k[x_i]/\frakm$. This gives us a surjection of $k[x_1,\dots,x_n]\to L$. Thus $\dim_k L$ is countable!

@@

\section{October 4, 2019}
Today we are going to be talking a bit more about the existence of affine varieties. Max talked a bit about the philosophy of work in this course:
he made this extended metaphor concerning butterflies but the take-away is to take learning onto ourselves. :)

\subsection{Questions from last time}
\subsubsection{Maps and elements}
In the book we did this silly thing where we defined $\Spec A\eqdef\Hom_{\Alg}(A,k)$ and then idendified 
$A$ with $k[\Spec A]$ by $a(f)=f(a)$. This seems a bit silly at first, but it may have something to do with the fact that 
we are looking for a natural way to construct affine varieties without having to choose a presentation. We will hopefully see 
something about this by the end of the day.

\subsection{Back to the Nullstellensatz}
Recall that we defined the operators $Z$ and $I$ that ``do the work'' of the Nullstellensatz. We then wrote (cor.~\ref{nullstellensatz}) $I(Z(J))=\sqrt{J}$. The idea is that 
this will gives us the function structure on an affine variety.
\begin{prf}[Cor.~\ref{nullstellensatz}]
	One way is not too hard. For the more difficult direction:
	Let $g\in I(Z(J))$. Then $Z(J)\subseteq Z(g)$. Now notice taht $D(g)$ can be naturally identified 
	with $\Spec k[x_i][1/g]$. Then consider
	\[J'=Jk[x_i][1/g]\]
	and the key realization is that $J'$ cannot be contained in any maximal ideal. The idea is that you can work by contradiction:
	this implies that $J$ is contained in an element of $D(g)$, but it isn't!

	Thus $J'=(1)$. So we can write $1=\frac{f}{g^N}$. Thus $g^k(f-g^N)=0$ in $k[x_i]$ and since $g$ isn't nilpotent, $f=g^N$.
\end{prf}
\begin{cor}
	There is a lattice anti-isomorphism between the radiacal ideals in $k[x_i]$ and Zariski-closed 
	subsets $Z\subseteq k^n$ via the maps $J\to Z(J)$ and $Z\mapsto I(Z)$.
\end{cor}
\begin{cor}
	For any ideal $J\subseteq k[x_i]$, 
	\[\sqrt{J}=\bigcap_{\text{maximal } \frakm\supset J}\frakm\]
\end{cor}
\begin{rmk}
	``The functions that vanish at the zero locus of $J$ are precisely those that vanish at all the points of $J$''.
\end{rmk}
\begin{cor}
	$D(g)\subseteq k^n$. Then the map 
	\[k[x_i][1/g]\to \Hom(D(g),k)\]
	via the map 
	\[\frac{f}{g^N}\mapsto \left(x\mapsto \frac{f(x)}{g(x)^n}\right)\]
	is injective.
\end{cor}

\subsection{Affine space}
Let's define $\A n\eqdef k^n$ with the Zariski topology. Let 
\[\calO_{\A n}(U)=\{f\in k(x_1,\dots,x_n)|\operatorname{poles}(f)\subseteq\A n\setminus U\}\subseteq\Hom(U,k)\]
Then, for example,
\[\O_{\A n}(D(g))=k[x_1,\dots,x_n][1/g].\]
\begin{prop}
	$\A n$ is an affine variety.
\end{prop}
\begin{prf}
	$\phi:X\to \A n$ gives us maps $\phi_1,\dots,\phi_n:X\to k$. Then that $\A n$ is affine relies on the fact:
	$\phi$ is a morphis if and only if the $\phi_i$ are regular. One direction is not too bad since coordinate functions 
	are regular by the axioms of morphisms. The other direction needs to be completed! DO IT!
\end{prf}

### 2019-10-04 Algebraic Groups/Algebraic-Groups.tex
Update

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Spec}{Spec}

@@

>>>>

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Spec}{Spec}
\newcommand{\red}[1]{{#1}_{\text{red.}}}

@@

\section{October 4, 2019}
Some review. Let $G$ be an algebraic group and denote $e:\Spec k\to G$ be the identity. We saw a lot of propositions last time.

Now let $k$ be a nonperfect field and take $t\in k\setminus k^p$. Then define 
\[G\eqdef V(x^{p^2}-tx^p)\subseteq \bbG_a\]
which Milne claims is not reduced. We can see why it's not geometrically reduced, but we're missing the details here.

\subsection{Another special case}
\begin{thm}
	When $k=\bar k$, $G$ is smooth if and only if 
	\[\dim T_e \red G=\dim T_e G.\]
\end{thm}
\begin{rmk}
	When $G$ is smooth, it is reduced, so the equality is clear. For the other direction, we get that $k$ is pefect, so $\red G$
	which is geoemetrically reduced if and only if $G$ is smooth. But 
	\[\dim G\le \dim T_e G=\dim T_e\red G=\dim \red G=\dim G\]
\end{rmk}
\begin{thm}
	If $G$ is a linear algebraic group over $k$ and $\ch k=0$, $G$ is smooth,
\end{thm}
\begin{prf}
	We can assume $k=\bar k$. Then set $G=\Spec A$ where $A$ is a Hopf algebra. THen we get Hopf algebra maps $m^\ast$ and $e^\ast$.
	Notice that the augmentation ideal $\frakm=\ker(e^\ast)$ is a maximal ideal.

	Then we want to prove 
	\begin{enumerate}
		\item $A\cong\frakm\oplus k$ as a $k$-vector space (obvious).
		\item $\forall a\in\frakm$, $m^\ast(a)-a\otimes 1-1\otimes a\in \frakm\otimes \frakm.$
	\end{enumerate}
	To see the second, notice that $m^\ast(a)-a\otimes 1-1\otimes a$ is in the kernel of 
	\[e^\ast\otimes \id:A\otimes A\to k\otimes A.\]
	This is clear from the commutative diagram
	\begin{center}
		\begin{tikzcd}
			k\otimes A & A\otimes A\ar[l,"e^\ast\otimes\id"] \\
			& A\ar[ul,"\sim"]\ar[u,"m^\ast"]
		\end{tikzcd}
	\end{center}
	Then we conclude $f\in \ker(e^\ast\otimes\id)\cap\ker(\id\otimes e^\ast)=A\otimes\frakm\cap\frakm\otimes A$ by a symmetric argument.
	Finally we notice that $A\otimes\frakm\cap\frakm\otimes A=\frakm\otimes\frakm$, and so $f$ lies in this ideal.

	Now we want to show that $\dim T_e G=\dim_k\frakm/\frakm^2=\dim_k\frakm/(\sqrt{0}+\frakm^2)=\dim T_e\red G$. It suffices to show that 
	for all $a\in\sqrt{0}$, $a\in \frakm^2$. Suppose the opposite--so let $a\in\sqrt{0}\setminus \frakm^2$. Consider the diagram 
	\begin{center}
		\begin{tikzcd}
			A\ar[r]\ar[d] &A_\frakm\ar[d]\\
			A/\frakm^2\ar[r,"\sim"] & A_\frakm/(\frakm A_\frakm)^2
		\end{tikzcd}
	\end{center}
	Now the image of $a$ in $A_\frakm$ is nonzero, so there exists $n>0$ such that $a^n\in A_\frakm$ but $a^{n-1}\ne 0$ in $A_\frakm$.
	Thus there exists $f\notin \frakm$ such that $a^nf=0\in A$. Subsititute $af$ for $a$, and thus there is an $a\in\sqrt{0}$ such that $a^n=0$ in $A$ but $a^{n-1}\ne 0$ in $A_\frakm$.

	Then by fact 2,
	\[m^\ast(a)=1\otimes a+a\otimes 1+r,\quad r\in\frakm\otimes\frakm\]
	and since $m^\ast$ is a ring homomorphism, 
	\[0=m^\ast(a^n)=(m^\ast(a))^n=(a\otimes 1+(1\otimes a+r))^n=a^n\otimes 1+n(a^{n-1}\otimes a+(a^{n-1}\otimes 1)r)+X\]
	where $X\in A\otimes\frakm^2$. But since $a^n=0$, we get 
	\[n(a^{n-1}\otimes a+(a^{n-1}\otimes 1)r)\in A\otimes \frakm^2\]

	Now since $(a^{n-1}\otimes 1)r\in (a^{n-1})\frakm\otimes A$, so 
	\[n(a^{n-1}\otimes a)\in (a^{n-1})\frakm\otimes A+A\otimes\frakm^2\]
	and since $\ch k=0$, we get that $n$ is a unit, so 
	\[a^{n-1}\otimes a\in (a^{n-1})\frakm\otimes A+A\otimes\frakm^2\]

	Now since this lives in $A\otimes A$, consider the image of the quotient map $A\otimes A\to A\otimes A/\frakm^2$. Then 
	\[a^{n-1}\otimes\bar a\in (a^{n-1})\frakm\otimes A/\frakm^2\subseteq A\otimes A/\frakm^2\]
	And note that $a^{n-1}\notin a^{n-1}\frakm$ because otherwise $a^{n-1}=a^{n-1}q$ for $q\in \frakm$.
	Then $a^{n-1}(1-q)=0\in A_\frakm$, which implies that $a^{n-1}=0\in A_\frakm$ (since $1-q$ is a unit here).

	Then somehow we get that $\bar a=0\in A/\frakm^2$, so $a\in \frakm^2$.
\end{prf}
